### 2022-02-22 manuscript_1.0/manuscript_1.0.Rmd
Error found in the manual coding of the cleaner entries used to compute visitor preference. In loadFieldData.R a programatic way of coding cleaner entries was done. Now that the manual and programatic results coincide I can re-run the chains. New figure configurations were implemented. Among those supplementary figures showing the diagnostics of the MCMC. A new result section was written.

---
title: Combination of field and experimental data together with computational models 
  reveal cognitive mechanisms behind cleaner fish behaviour
#Cognitive and ecological aspects determine cleaner performance
author:
  - name: Andrés E. Quiñones 
    affiliation: "1"
  - name: Zegni Triki
    affiliation: "2"
  - name: Redouan Bshary
    affiliation: "1"
address:
  - code: "1"
    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - code: "2"
    address: Department of Zoology, Stockholm University, Stockholm, Sweden

corresp_author_name:  "AE. Quiñones"
corresp_author_email: "andreseqp@gmail.com"

subject:
  - "Behavioural ecology"
  - "Cognitive ecology"
  - "Animal behaviour"

keywords:
  - learning   
  - cleaners
  - behaviour
  - bayesian statisitics
  - mutualism

abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner fish relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance.

## Remove this if not required
ethics: |
  The Animal Ethics Committee of the Queensland government (DAFF)
  approved the project (CA 2016/05/970 and CA 2017/05/1063).

data_accessibility: |
  Please provide details on the data availability.

author_contributions: |
  AQ and RB designed the study. ZT collected the data. AQ developed the model and 
  fitted the model parameters to the data. AQ, RB and ZT wrote the manuscript.

## Remove this if not required
conflict_of_interest: |
  The authors declate no conflict of interests.

## Remove this if not required
funding: |
  This work was supported by the Swiss National Science Foundation (grants
  number: 31003A_153067/1 and 310030B_173334/1 to R.B.).

## Remove this if not required
acknowledgements: |
  ZT and RB kindly thank the staff of Lizard Island Research Station, Dominique
  Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena Levorato and William McNeely for their
  field support. AQ thanks Florian Hartig for statistical advice.

bibliography: Cleanerlearning.bib

## change to true to add optional line numbering
lineno: true

header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{booktabs}
  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
site: bookdown::bookdown_site
output: 
  bookdown::pdf_book:
    base_format: rticles::rsos_article
---

## Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
source(file=here("loadFieldData.R"))
```

Disentangling the mechanistic underpinnings of behaviour 
is typically done under controlled conditions in a laboratory setting. However, 
controlled laboratory conditions often mask the inter and intra-specific 
variation that arises in the interaction between mechanisms and 
environmental variation. From a evolutionary perspective, mechanisms 
are likely selected because of how they allow individuals to respond to 
environmental variation. For example, biological market theory predicts that 
the exchange rate of goods and/or services traded between cooperative 
partners adjusts to the law of supply and demand, when individuals have 
some degree of partner choice [@noe_Biological_1995a]. 
Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station at their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish had eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. Given that primates 
are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, this raises the question:
what makes the task difficult for most non-cleaners?

Recent research on the cognitive tool kit needed to solve the ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates 
 [@wismer_Cuebased_2019] while others find it in the food 
[@pretot_Comparative_2021; @pretot_Comparing_2016]. However, identifying 
cues as salient is not the only challenge of the ephemeral reward task, 
as revealed by proximate learning models. Such models allow varying the 
cognitive tool kit and to evaluate which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value differently
one stimulus alone from compound versions of it has been termed 
configurational learning, or chunking, 
or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is in principle 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, the two mechanisms should 
be evaluated in light of how well they explain the available data.

Over the last decade, our team tested over a hundred of wild-caught 
cleaner fish in the exact same paradigm of  of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaners should not give priority to visitors any more.

The interplay between the underling cognitive machinery and 
local ecological conditions apparently generate the documented variable 
performance among individuals of the same species. 
Our approach of fitting the computational model to the empirical data on 
fish densities and cleaner fish censuses and performance
in the ephemeral reward task aimed at: (i), determining which 
mechanism cleaner fish use to incorporate future consequences of current 
decisions by testing whether operant conditioning, chaining, or a combination 
of both best explains their performance; (ii)  determining whether the 
two mechanisms differed with respect to the ecological conditions 
that are likely to cause high versus low performance in the 
ephemeral reward task. Additionally, we assessed which mechanism 
yields optimal performance patterns. Relying on the logic of 
biological market theory, we predicted 
that appropriate performance is to show low preference for visitors under low
local cleaner-to-client ratio. 
<!-- In this case, as visitor leaving rates  -->
<!-- should be low as well and any leaving visitor will be readily  -->
<!-- replaced by another client in the queue. In contrast,  -->
<!-- cleaners should show high performance in the ephemeral reward task -->
<!-- if the local cleaner-to-client ratio favours clients.  -->

## Methods

<!-- ### Overview -->

<!-- In this study, we combined empirical data from Triki *et. al*  -->
<!-- [@triki_Biological_2019;@triki_Brain_2020] and the computational  -->
<!-- model analysed in Quiñones *et al.*[@quinones_Reinforcement_2019]  -->
<!-- to test the relative validity of chaining versus operant conditioning as  -->
<!-- mechanisms to account for future consequences of cleaner fish desicion-making. -->
<!-- In order to compare the explanatory power of the different hypotheses,  -->
<!-- we estimated the parameter values of three versions of  -->
<!-- the model that best fit the empirical data.  -->
<!-- Particularly, we fitted the parameter that determines how important  -->
<!-- future reward is ($\gamma$), the magnitude of the negative reward  -->
<!-- ($\eta$) and a scaling constant that translates cleaner fish absolute  -->
<!-- to relative population abundances (see following sections for details).  -->
<!-- We obtained the model predictions using a set of stochastic individual-based -->
<!-- simulations. Thus, each parameter combination of the model yielded not one  -->
<!-- but a distribution of predictions. For the  -->
<!-- fitting procedure, we used the bayesian inference algorithm:  -->
<!-- Markov Chain Monte Carlo (MCMC). MCMC is a stochastic algorithm that allows -->
<!-- to sample from the posterior distribution of model parameter values. The posterior -->
<!-- distribution represents the most likely parameter values given the data.  -->
<!-- Due to the stochastic nature of MCMC, which repeatedly samples parameter values -->
<!-- and evaluates their likelihood, the distribution of likelihoods from the  -->
<!-- stochastic model is also explored. This makes MCMC and bayesian inference -->
<!-- particularly suitable for the computational models we evaluate. Bellow,  -->
<!-- we explain first the computational model and the data collection, and later  -->
<!-- how the MCMC was used to estimate parameter values. -->

### The model

The model consists of a set individual-based simulations where
individuals, representing cleaners, face a series of choices 
between two options, which simulate the natural conditions 
of the cleaning market. Individuals experience a series 
of discrete time points in which they face different ‘states’, 
defined by the number and category of client fish (visitor or residents) 
inviting for cleaning services. There are six possible states: 
zero clients, one resident, one visitor, 
resident-resident, visitor-visitor, and resident-visitor. 
The probability of each state is largely determined by the relative 
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination. 
This is because residents are willing to queue for cleaning service; 
while visitors leave the queue (with a certain probability) when made to wait. 
Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expected reward and the obtained
reward (prediction error) [@sutton_Reinforcement_2018; 
@rescorla_Theory_1972]. Given the new information, the update is carried
in the direction that leads to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor 
over a resident converges in the model. To which probability the model 
will converge depends on the relative abundance of cleaners, visitors 
and resident; as well as on the probability of visitors leaving the 
cleaning station when made to wait. Further details of the model 
implementation can be found in Quiñones *et al* [@quinones_Reinforcement_2019].

The model shows that agents need to find a mean 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

<!-- A further model analysis contrast the effect of these two  -->
<!-- parameters against variation in the relative abundance of cleaners  -->
<!-- and the two client types. Assuming certainty that visitors leave, the  -->
<!-- two parameters yield different  -->
<!-- predictions when relative client abundance is very high. Under such  -->
<!-- conditions, the analysis shows that under very high client (relative)  -->
<!-- abundance the presence of future rewards ($\gamma>0$) no longer favours  -->
<!-- the preference for the ephemeral clients, while the preference persists  -->
<!-- when the negative reward mediated by client behaviour ($\eta>0$) is  -->
<!-- included. Thus, the analysis suggests that future reward as an  -->
<!-- explanation for cleaner fish ability to prefer visitors is supported.  -->
<!-- At first sight, empirical data seem to support future rewards as  -->
<!-- the mechanism used by cleaners, as the statistical analysis  by  -->
<!-- Triki *et al.* [@triki_Biological_2019] data revealed that low  -->
<!-- cleaner-to-client ratios predict low preference for visitors and hence  -->
<!-- low performance in the ephemeral reward task. However, low cleaner-to-client  -->
<!-- ratios may also lead to low probabilities of visitors leaving, which would  -->
<!-- cause low preferences for visitors independently of the underlying  -->
<!-- mechanism [@quinones_Reinforcement_2019]. Furthermore, a combination  -->
<!-- of future and negative rewards, could still be yield a better explanation  -->
<!-- for the data than any single factor.  -->

### Empirical data

The empirical data were collected between 2010 and 2019 always during 
the austral winter months June to August from a total of five 
study reef sites (Corner Beach-CB, Horseshoe-HS, Mermaid Cove-MC, 
Northern Horseshoe-NHS, and The Crest-TC)  at Lizard Island 
($14.6682° S, 145.4604° E$), Great Barrier Reef, Australia. 
The data consist of three sets: fish censuses, field 
observations of cleaner-client interactions to quantify the probability 
of visitors leaving if made to wait, and the 
performance of wild-caught cleaner fish in the ephemeral reward test. 
In total, we have twelve site/year data sets for fish censuses and 
corresponding performance in the lab test. Thus, some sites were sampled more 
than once. To estimate the population density of cleaner fish 
and their clients at a given site in 
a given year, Triki *et al.* [@triki_Biological_2019] used a series of ten 
transects of $30m$ each. Observers swam along the transect lines placed on 
the reef and first counted the visible large-bodied adult fish 
(species with total length $TL \geq 10cm$) including cleaner fish 
on a width of $5m$, and then on the 
return individuals of small-bodied fish species ($TL < 10 cm$) \
on a width of $1 m$
(see Triki et al. 2018 for further details on fish censuses data collection). 
We then scaled the counts of cleaner fish, small-bodied, 
and  large-bodied  clients fish densities per $100 m^2$. 
From the study by Triki *et al* [@triki_Biological_2019] 
cleaner fish and large-bodied client populations densities were 
highly correlated, and only the former 
was hence used in the analyses as representative of both measures.

<!-- Cleaner densities, in interaction with visitor client leaving probability,  -->
<!-- explained best individual performance in the ephemeral reward task  -->
<!-- [@triki_Biological_2019], while cleaner-to-client ratio was not a significant  -->
<!-- factor. This creates a certain mismatch with the model by Quiñones *et al.*  -->
<!-- [@quinones_Reinforcement_2019], which focused on abundance of cleaners relative  -->
<!-- to overall client abundance in order to capture the characteristics  -->
<!-- of the market. Importantly, relative abundances of clients define not only  -->
<!-- the probability of interactions with residents and visitors, but also affect  -->
<!-- how often the cleaning station is empty (e.g. there are no clients to be cleaned).  -->
<!-- The higher the relative abundance of cleaners is, the higher is the probability  -->
<!-- of cleaning stations being empty at a given time point. How often the  -->
<!-- cleaning stations are empty in turns influences the value that cleaners can  -->
<!-- expect from their choices. The emptier the cleaning stations are, the more  -->
<!-- beneficial it is to prioritize a client that will not stay. On the other hand,  -->
<!-- situations where cleaners have to choose will be rare if cleaner fidh relative  -->
<!-- abundance is high, which reduces the total value of prioritizing visitors  -->
<!-- over residents and makes it more difficult to learn a preference for visitors.  -->
<!-- Given the model structure, we used the empirical data to calculate relative  -->
<!-- abundances of cleaners, residents and visitors for the current analyses.  -->
<!-- To incorporate the empirical finding that absolute densities matter, we  -->
<!-- introduced a new scaling constant (see section on model fitting).  -->

The field observation data consisted of video recordings/encodings of the 
cleaner-client cleaning interactions. There were videos from eight 
cleaners per site/year of a duration of $30 min$ each. Triki *et al* 
extracted information from every event wherein a visitor client 
was made to wait in favour of another client (visitor or resident), and noted 
whether or not the visitor left or queued for the cleaning service
[@triki_Biological_2019;@triki_Brain_2020].

The cognitive performance data was from a total of 120 cleaners 
(10 individuals per 12 site/year) tested in the 
ephemeral reward task [@triki_Biological_2019;@triki_Brain_2020]. Authors 
housed all captured cleaners individually in glass aquaria 
( $62cm \times 27cm \times 37 cm$ ) and provided them 
with PVC pipes ($10 cm \times 1 cm$) as shelters. 
The task consisted of exposing the cleaner fish to substitute 
models of client fish in the form of two *Plexiglas* plates offering the 
same amount of food (one item of mashed prawn).  The two plates differed 
in colour and pattern (horizontal green stripes or vertical pink stripes) 
but had equal size ($10 cm \times 7 cm$). Importantly, the two plates played 
different roles as either a visitor (ephemeral food source) or 
resident (permanent food source). That is, if a cleaner fish inspected the 
resident plate first, the experimenter withdrew the visitor plate out of 
the aquarium as a consequence.  Choosing first the visitor plate, 
however, granted access to both plates. The equal size of the plates 
forced cleaner fish to learn to give service priority to the visitor plate 
based solely on the behaviour-cue of the plates rather than size-cue 
[@wismer_Cuebased_2019]. Triki *et al.* [@triki_Biological_2019;@triki_Brain_2020] 
tested the fish for a maximum of 200 trials with 20 trials a day, 10 trials 
in the morning and 10 trials in the afternoon. They randomized and 
counterbalanced the plates’ spatial location (i.e. left or right) 
between trials. Similarly, they counterbalanced the plates’ decoration 
(colour and pattern) and the plates’ role (visitor or resident) between the 
tested fish. In the original studies, once a fish reach a learning criterion, 
that is, performing significantly above chance level ($> 50%$, $p-value ≥ 0.05$),
they passed to a reversal version of the task where the roles of the 
visitor/resident Plexiglas plates were swapped (see Triki et al. 2019, 2020). 
Here, we used instead a subset of these data in order to have an idea on 
cleaner fish preferences for the visitor plate, even if they do not reach 
the learning criterion within 200 trials. To do so, we first extracted the 
trial-by-trial outcomes from the last two session (20 trials) of those who 
never reached the learning criterion for visitor plate 
($N = `r fieldData.cleaner.filt.sum[V1==2,length(V1)]`$ cleaner fish). 
For those who reached the learning criterion at some point during the 
test and passed to a reversal phase, we extracted the trials-by-trial 
outcomes from the last session (10 trials) before passing to reversal 
and the last 10 trials they were exposed to in the test 
($N = `r fieldData.cleaner.filt.sum[V1>2,length(V1)]`$ cleaner fish).

### Statistical analysis

The aim of the analyses is to fit the key model parameters 
$\gamma$ and $\eta$, to the empirical data from Triki *et al.* 
[@triki_Biological_2019;@triki_Brain_2020] to test whether each or 
a combination of these effects is a better explanation for the pattern 
seen in the data. We used the ecological variables: cleaners, visitor clients, 
resident clients abundances and visitor clients leaving probability,  
as input to the models. As response variable we used cleaners' 
preference for visitor over residents in the ephemeral reward task. 
Finally, we used the preference for visitors resulting from the 
model simulations as the prediction for the response variable. 
We kept all other parameter values used for the model simulations constant, 
see Table \ref{tab:param}.

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors we need to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because in the model, relative abundances of clients 
define not only the probability of residents and visitors, 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable, which in nature may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equals to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used a 
uninformative prior. Once we decided whether the new parameter set replaces 
the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters, and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
$\color{red}{\text{include citation}}$

To compare the fit of the three alternative models, we used the 
distribution of pseudo $R^2$ proposed by Mcfadden [@mcfadden_Conditional_1974].
Mcfadden's pseudo $R^2$ is a standard measure of fit for logistic regression. 
In that context, $pseudo-R^2$ uses the log-likelihood of the data given 
the model, relative to the log-likelihood of the data given a 
model without covariates, as a measure of fit. Our 
model is not a logistic regression, therefore we measured the $pseudo-R^2$ in
relation to the log-likelihood of a model
with parameters $\gamma$ and $\eta$ set to zero. This in practice
amounts to a model that has neutral preference between the two options. 
We computed the pseudo $R^2$ for all the samples of the posterior from
the MCMC. Thus, we used these distributions of pseudo $R^2$'s as a measure of
fit.

## Results

(ref:post) Posterior distributions for parameter values $\gamma$ and $\eta$ for 
the three model ("Full model", "chaining only" and "penalty only"). We show 
the kernel density estimates, below the mode (black dot) and the 65% 
(light blue shade) and 95% (grey shade)  highest posterior density interval 
for the two parameters. On the top, panels a and b show posterior 
distributions for a full model, including chaining ($\gamma$) and penalty 
($\eta$). Panel c, shows the $\gamma$ estimate from model with only chaining. 
Panel d shows the $\\eta$ estimate from a model with only penalty. Panel e, 
shows a measure of fit for all models, namely the distribution of pseudo-$R^2$ 
obtained from sampling the posterior distribution of parameter values.

```{r post, fig.cap='(ref:post)',out.width='100%',fig.height=6}
library(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
library(ggpubr)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

# Name of the folders containing the MCMC runs
scen1<-"MCMCclean_gam_nRew_sca_"
scen2<-"MCMCclean_gam_sca_"
scen3<-"MCMCclean_Nrew_sca_"

# NLables names for the three scenarios
labels.Scen<-c("both","gam","Nrew")
labelsPlot.Scen<-c("Full Model","Chaining","Penalty")

## Load files --------------------------------------------------------------

# Defaults for MCMC analysis
burn.in<-1000
thinning<-100

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
# MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

# Create list with the data, convenient to use CODA for diagnostics
mcmcList.1<-mcmc.list(do.call(list,lapply(unique(MCMCdata1$seed), function(repli){
  rundata<-MCMCdata1[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)
  return(mcmcRun)
})))

# mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
#   rundata<-MCMCdata3[seed==repli]
#   mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)#
#   return(mcmcRun)
# })))

## Density plots with gg -------------------------------------------------------

## Define the boundaries for credible intervals using High Density Interval

intervals<-c(0.66,0.95)

cuts.1<-lapply(MCMCdata1[,.(gamma,negReward,scaleConst)],
             function(x){
               c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
                 mode_hdci(x,.width = intervals)$ymax)
})

cuts.2<-lapply(MCMCdata2[,.(gamma,scaleConst)],
               function(x){
                 c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdci(x,.width = intervals)$ymax)
               })

# cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
#                function(x){
#                  c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
#                    mode_hdci(x,.width = intervals)$ymax)
#                })

# get all the likelihoods from the posterior for all models
# loglikehoods.all<-data.table(lglikelihood=c(MCMCdata1$fit,MCMCdata2$fit,
#                                              MCMCdata3$fit),
#                               model=c(rep(labels.Scen[1],length(MCMCdata1$fit)),
#                                       rep(labels.Scen[2],length(MCMCdata2$fit)),
#                                       rep(labels.Scen[3],length(MCMCdata3$fit))
#                                       ))
# 
# # Read field data to calculate null log-likelihood
# fieldData<-fread(here("Data","data_cleaner_abs_threa1.5.txt"))
# 
# nullLikehood<-sum(dbinom(x=fieldData[,score_visitor],
#                          size = 20,prob = 0.5,log = TRUE))
# 
# # calculate pseudo R^2
# loglikehoods.all[,Rsqrd:=1-lglikelihood/nullLikehood]
# 
# 
# panel.rsqr.all<-
#   ggplot(data=loglikehoods.all[is.finite(lglikelihood)],
#                   aes(y=model,x=Rsqrd,fill=model))+
#   stat_halfeye(alpha=0.5,point_size=3)+
#   theme_classic()+#xlim(-0.5,0.25)+
#   scale_fill_manual(values=c("#1b9e77","#d95f02","#7570b3"),
#   name = "Model", labels = labelsPlot.Scen)+
#   scale_y_discrete(labels=labelsPlot.Scen)+
#   geom_vline(xintercept = 0,color="black",size=0.5)+
#   theme(legend.position = 'none')
#           # c(.32, .65)
#         # ,legend.key.size = unit(0.22,'cm'),
#         # axis.text.y = element_blank(),
#         # legend.text = element_text(size=7))
# 
# gam.both.post<-ggplot(data=MCMCdata1,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,show.legend=FALSE) +
#   labs(title=labelsPlot.Scen[1],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))
# 
# nrew.both.post<-ggplot(data=MCMCdata1,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$negReward))),
#                point_interval = mode_hdi,
#                .width = c(.66, .95),point_size=3,
#                show.legend=FALSE)+theme_classic()+
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title="",subtitle = expression(eta),x="",y="")+
#   xlim(-2,2)
# gam.gam.post<-ggplot(data=MCMCdata2,aes(x=gamma)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$gamma))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[2],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()
# nrew.nrew.post<-ggplot(data=MCMCdata3,aes(x=negReward)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$negReward))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[3],
#        subtitle = expression(gamma),x="",y="")+
#   theme_classic()+xlim(-2,2)
# 
# first.col<-plot_grid(nrow=3,align = "v",byrow = TRUE,
#           gam.both.post,gam.gam.post,nrew.nrew.post,
#           labels=c('a','c','d'))
# second.col<-plot_grid(nrow = 2,align = "v",byrow = TRUE,
#           nrew.both.post,panel.rsqr.all,rel_heights = c(1,2),
#           labels = c('b','e'))
# plot_grid(ncol = 2,first.col,second.col)

```

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centred around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models` fit 
favours the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though, the peak of the three
$pseudo-R^2$ distributions was not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of black line in Fig. \ref{fig:post} f). This
shows that, accounting for variation in the parameter estimates, the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions, given that they do not bring 
biological insight. Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagneg}).

(ref:pred) Observed and predicted probability of choosing a visitor. 
Left-hand side panel: colour contour shows the prediction from the 
learning model using the mode of the posterior distributions of parameters 
recovered by the statistical analysis. Dots show the frequency of visitor 
choices for the 12 reef sites, as well as the corresponding relative 
cleaner fish abundance (x axis) and frequency of visitors clients leaving 
the cleaning station (y axis). Right-hand side panels: Variation of the 
predicted probabilities  of choosing a visitor over a resident and their 
observed values for 12 locations. Circles show the mean prediction for 
each location from 100 samples taken from the posterior distribution. 
Thick and thin bars show the 66 and 95% credible interval, respectively, 
taken from those posterior samples. Squares show the predictions used for 
panel on the right-hand side. Colour coding denotes different reef sites/year 
of the data collection (see Empirical data section). Black line corresponds 
to a perfect match between observed and predicted probabilities. Upper 
panels (a and b) show predictions from a model including chaining and 
penalty; middle panels (c and d) from a  model with only chaining; lower 
panel (e and f) from a model with only penalty.

<!-- # ```{r pred, fig.cap='(ref:pred)',out.width='100%',fig.height=5} -->
<!-- #  -->
<!-- # source(here("manFigures.R")) -->
<!-- # ggarrange(cont.obs.pred.both,scatter.obs.pred.both, -->
<!-- # cont.obs.pred.gam,scatter.obs.pred.gam, -->
<!-- # cont.obs.pred.Nrew,scatter.obs.pred.Nrew, -->
<!-- #           labels=c('a','b','c','d'),common.legend = F, -->
<!-- #   heights=c(1.25,1)) -->
<!-- #  -->
<!-- # ``` -->

The main reason for chaining and penalty to give different
predictions is the way that cleaner fish relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models, they all predict an increase preference 
for visitors as the visitor probability increases  (Fig. \ref{fig:pred}).
In contrast, cleaner fish relative abundance has a different effect in the
model with only chaining, compared to the other two models,
full model and penalty. In the model with only chaining, only intermediate 
cleaner fish abundance triggers a preference for the visitor 
clients (Fig. \ref{fig:pred} c). In both models with penalty, 
both intermediate and low cleaner abundance triggered a preference 
for the visitor (Fig. \ref{fig:pred} a,e). Note, however, we calculated 
preferences shown in figure \ref{fig:pred} left panels by using only the mode 
of the posterior distributions, and by holding constant the balance between 
resident and visitors´s abundances. Panels on the right, show how close 
predictions are from the observed data, allowing the balance between 
client types to vary, and using a set of samples from the posterior 
distribution.

## Discussion

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, operant conditioning, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the these fish apply 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients, and generalized this rule to the lab conditions 
once being familiar with the task.

While all three models captured well the positive relation between visitor 
leaving behaviour and cleaner fish performance in
the market task [@triki_Biological_2019], only the chaining mechanism 
predicted that cleaner fish performance in the task should be low in 
habitats with low cleaner-to-client ratios, regardless of the 
visitor leaving probability. In contrast, models including 
negative reward predicted the highest performance in the ephemeral 
reward task when relative cleaner fish abundance is low, particularly 
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low relative cleaner fish abundances mean the market has an excess of demand for
cleaning services. In the models, this translates to a cleaning station
that is frequently full. Thus, when a visitors leaves, it is likely that the
cleaner fish will have access to another client in the next step. Therefore, 
there will not be much difference in future reward between choosing 
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward 
on cleaner fish preference is the opposite, as in a busy cleaning station, to
that of chaining. cleaner fish will get more often the resident-visitor
state and will develop a preference for the visitor faster. 
At high cleaner fish abundances, the resident-visitor state
becomes so rare that neither mechanism is very efficient 
at generating a preference for visitors. When facing the resident 
visitor choice it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, the models suggest that chaining is the cognitive
mechanism that allows cleaner fish to adaptively adjust to their 
biological market ecological conditions.

Previous research showed that cleaner fish living at high population 
densities and giving service priority to the visitor plate in the 
ephemeral reward task, as well as cleaner fish living at low 
densities but denying service priority to the visitor plate possess 
larger forebrains; a key teleost brain region associated with 
behavioural flexibility and social intelligence. Those failing to 
show optimized decision-rules given their local ecological 
conditions had relatively smaller forebrains  [@triki_Brain_2020]. 
Triki *et al.*  refer to the former as socially competent cleaner fish,
while the second group as socially incompetent cleaner fish. 
Social competence is the ability to optimise social behaviour 
to the available social information [@taborsky_Social_2012a;
@bshary_Cooperation_2015;@varela_Correlated_2020]. Our analyses yielded no 
evidence that the difference in social competence with respect to the local 
ecological conditions and associated brain morphology, 
found by Triki *et al* [@triki_Brain_2020], are due to the mechanism 
used to incorporate future consequences. It could have been conceivable that 
high performing individuals from low population densities reef sites 
might use negative reinforcement instead of chaining, but in that 
case negative reinforcement should have explained at least 
part of the data. Configurational learning or 
chunking [@sutherland_Configural_1989;@miller_Magical_1956], 
the second component necessary to solve the ephemeral 
reward task [@quinones_Reinforcement_2019], was 
not varied in the models we analized here. However, while chunking tendencies 
should vary to allow individuals to adapt to their local conditions 
[@prat_Modelling_2022;@kolodny_Evolution_2014], systematic differences 
in individual chunking tendencies would not explain how socially 
competent decisions vary as a function of relative abundance. 
Therefore, it remains currently unclear what cell-demanding 
mechanisms may cause variation in social competence that 
translates into site-specific variation in performance in the 
ephemeral reward task.

Our models are inspired by the general processes of associative learning
where short term rewards are translated into decision making; thus, it 
ignores alternative channels of information that could be relevant in 
market-like situations. For example, the model does not investigate whether 
cleaner fish actually assess the frequency of client visits or a mean frequency 
of visitors leaving. The updating learning mechanism for the development 
of preferences works on a trial-by-trial basis. In the model, cleaner fish 
do not need to assess the actual state of the market, *i.e.* their abundance, 
the abundance of residents and visitors, and client visitation rate as an 
indicator of demand. They only need to assess short-term consequences of their
own decisions on food intake, and chain them. Also, for the sake of 
simplicity, the model ignores the process by which cleaner fish discriminate 
residents and visitor clients. A model that accounts for this discrimination
probably would involve the development of preferences for morphological or 
behavioural features that are statistically associated with visitors 
or residents. For example, visitors are on average larger 
than residents in body size [@bshary_Cleaner_2001a], and contrary to 
resindents, are less likely to chase a cleaner fish that fails to cooperate
and intead cheats its client by taking a bite of mucus 
[@bshary_Asymmetric_2002]. Given these associations, chaining might 
produce the decision-rule “choose the larger client and/or the less aggressive 
client”, which is not a useful rule in the standard ephemeral reward task.

In conclusion, our study shows that variation in cognitive performance as 
a function of the local ecological conditions may set the stage for the use of 
mechanistic modelling to identify the cognitive processes underlying 
learning in animals. The combination overcomes the limitations 
of the general philosophy in animal cognition to apply the logic of 
Lloyd Morgen's canon (Occam's razor). Cognitive experiments with the 
aim of excluding basic reinforcement learning as a potential 
explanation (operant and/or classical conditioning) of performance 
often employ one trial experiments requiring animals to solve the 
task on the first possible occasion. 
For example, any theory of mind task needs to be solved in the first
trial in order to exclude fast conditioning [@heyes_Theory_1998].
Similarly, subjects need to solve a social learning task on the first
trial to accept imitation as a mechanism over stimulus/local
enhancement. Such strict conditions are virtually never met. For
example, potato washing by Japanese macaques, an iconic example of
social learning, took several years to spread within the group
[@kawamura_Process_1959], meaning that any learner had been repeatedly
exposed to demonstrations before acquisition. Importantly, Galef
[@galef_Question_1992] refuted imitation as mechanism not simply because
of the repeated exposure but because a (rather qualitative) analysis of
the spread of potato washing across individuals did not follow the
prediction based on imitation learning (see also
[@hirata_SweetPotato_2001]). In our case, the number of trials it took
cleaners to learn the solution to the ephemeral rewards task 
would never allow to exclude an important role of 
negative reinforcement based on the data
alone. However, fitting model predictions to our comprehensive empirical 
data set revealed that a more complex mechanism, estimation of future reward,
fits the data better.

\newpage

# Supplementary material

```{=tex}
\beginsupplement
```

(ref:rawdata) Relation between the response variable used in this study and the 
criteria used in previous studies to assess performance in the ephemeral reward 
task. In the x axis, we classified the performance of cleaner fish according to 
whether they developed a preference for the visitor in the initial round, 
in the initial and reversal, or none of them. In the y axis, we add the choices
of two experimental sessions: panel on the left uses one session from the initial 
round and one from the reversal round when possible 
(as described in the main text); panel on the right uses two sessions from the 
initial round for all fish.

```{r rawdata,fig.cap='(ref:rawdata)'}
plot_grid(initANDRev,init,ncol = 2)

```

(ref:scaConst) Posterior distributions for scaling constant for the three model 
("Full model", chaining only" and "penalty only"). We show the kernel 
density estimates, below the mode (black dot) and the 65% (light blue shade) 
and 95% (grey shade)  highest posterior density interval. On the top, 
panel a shows the posterior distribution from the full model; panel b from 
the model with only chaining; and panel c from a model with only penalty.

```{r scaConst, fig.cap='(ref:scaConst)',fig.pos = 'H', out.extra = "",out.width='95%',fig.height=4}

sca.both.post<-ggplot(data=MCMCdata1,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.1$scaleConst))),
               point_interval = mode_hdi, .width = c(.66, .95),
               point_size=3,show.legend=FALSE) + 
  labs(title=labelsPlot.Scen[1],x="",y="")+
  theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))
sca.gam.post<-ggplot(data=MCMCdata2,aes(x=scaleConst)) + 
  stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.2$scaleConst))),
               point_interval = mode_hdi, 
               .width = c(.66, .95),point_size=3,
               show.legend=FALSE)+theme_classic()+
  scale_fill_manual(values=c("gray85","skyblue","gray85"))+
  labs(title=labelsPlot.Scen[2],x="",y="")
# sca.nRew.post<-ggplot(data=MCMCdata3,aes(x=scaleConst)) +
#   stat_halfeye(aes(fill=stat(cut(x,breaks = cuts.3$scaleConst))),
#                point_interval = mode_hdi, .width = c(.66, .95),
#                point_size=3,
#                show.legend=FALSE) +
#   scale_fill_manual(values=c("gray85","skyblue","gray85"))+
#   labs(title=labelsPlot.Scen[3],x="",y="")+
#   theme_classic()
ggarrange(sca.both.post,sca.gam.post,
          # sca.nRew.post,
          labels=c('a','b','c'),common.legend = F)

```

```{r param , echo=FALSE,results='asis'}
  cat(' Table: \\label{tab:param} Parameter values with which the model was run 
  in the MCMC. $\\sigma$ refers to the amplitud of the perturbation kernel with the subscript indicating the associated parameter. New values were taken from a uniform distribution. $\\alpha$ refers to the learning rate.

  | Parameter | Value |
|----------:|:-----:|
| Learning rounds | 10000 |
| Reward value | 1 |
| $\\alpha$ | 0.05 |
| $\\sigma_{\\gamma}$ | 0.3 |
| $\\sigma_{\\eta}$ | 4 |
| $\\sigma_{Sca.Const.}$ | 300 |
| Number of chains | 5 |
| Chain lenght | $1^5$ |
')

```

(ref:diagnosticsfull) MCMC convergence diagnostics for the full model. 
On the left trace-plots, on the right changes along the chain of the 
Gelman and Rubin's shrink factor [@brooks_General_1998].

```{r diagnosticsfull, fig.cap='(ref:diagnosticsfull)',echo=FALSE,fig.height=5}

longMCMCdata1<-melt(MCMCdata1,id.vars = c("iteration","seed"),
                    variable.name = "parameter",
                    measure.vars = c("gamma","negReward","scaleConst"))

parlab<-list(`gamma` =expression(gamma),
          `negReward`=expression(eta),
          `scaleConst`="Scaling constant")

traces.full<-ggplot(longMCMCdata1,aes(x=iteration,y=value,colour=as.factor(seed)))+
  geom_line()+
  scale_color_manual(values = multDiscrPallet)+
  labs(color="Chain")+
  facet_grid(rows=vars(parameter),scales = "free",switch = "y",
             labeller = function(variable,value){
  return(parlab[value])})+
  theme_classic()+
  theme(legend.position = 'top')

gel.full<-as.data.table(gelman.diag(mcmcList.1)$psrf)
gel.full[,parameter:=c("gamma","negReward","scalConst")]

pdf(file = NULL)
gp.dat<-gelman.plot(mcmcList.1)
dev.off()

gp.dat.frame <- data.table(rbind(as.data.frame(gp.dat[["shrink"]][,,1]), 
                          as.data.frame(gp.dat[["shrink"]][,,2])), 
                q=rep(dimnames(gp.dat[["shrink"]])[[3]], each=nrow(gp.dat[["shrink"]][,,1])),
                last.iter=rep(gp.dat[["last.iter"]], length(gp.dat)))

gelman.full<-ggplot(melt(gp.dat.frame, c("q","last.iter"), value.name="shrink_factor"), 
       aes(last.iter, shrink_factor, colour=q, linetype=q)) + 
  geom_hline(yintercept=1, colour="grey30", lwd=0.2) +
  geom_line() +
  facet_grid(rows = vars(variable), labeller= labeller(function(var,value){
  return(parlab[value])})) +
  labs(x="Last Iteration in Chain", y="Shrink Factor",
       colour="Quantile", linetype="Quantile") +
  scale_linetype_manual(values=c(2,1))+
  theme_classic()+
  theme(legend.position = 'top')

ggarrange(traces.full,gelman.full)

```

(ref:diagnosticsGam) MCMC convergence diagnostics for the chaining model. 
On the left trace-plots, on the right changes along the chain of the 
Gelman and Rubin's shrink factor [@brooks_General_1998]

```{r diaggam,fig.cap='(ref:diagnosticsGam)',echo=FALSE}
longMCMCdata2<-melt(MCMCdata2,id.vars = c("iteration","seed"),
                    variable.name = "parameter",
                    measure.vars = c("gamma","scaleConst"))

parlab<-list(`gamma` =expression(gamma),
          `scaleConst`="Scaling constant")

traces.gam<-ggplot(longMCMCdata2,aes(x=iteration,y=value,colour=as.factor(seed)))+
  geom_line()+
  scale_color_manual(values = multDiscrPallet)+
  labs(color="Chain")+
  facet_grid(rows=vars(parameter),scales = "free",switch = "y",
             labeller = function(variable,value){
  return(parlab[value])})+
  theme_classic()+
  theme(legend.position = 'top')

gel.gam<-as.data.table(gelman.diag(mcmcList.2)$psrf)
gel.gam[,parameter:=c("gamma","scalConst")]

pdf(file = NULL)
gp.dat.gam<-gelman.plot(mcmcList.2)
dev.off()

gp.dat.frame.gam <- data.table(rbind(as.data.frame(gp.dat.gam[["shrink"]][,,1]), 
                          as.data.frame(gp.dat.gam[["shrink"]][,,2])), 
                q=rep(dimnames(gp.dat.gam[["shrink"]])[[3]], each=nrow(gp.dat.gam[["shrink"]][,,1])),
                last.iter=rep(gp.dat.gam[["last.iter"]], length(gp.dat.gam)))

gelman.gam<-ggplot(melt(gp.dat.frame.gam, c("q","last.iter"), value.name="shrink_factor"), 
       aes(last.iter, shrink_factor, colour=q, linetype=q)) + 
  geom_hline(yintercept=1, colour="grey30", lwd=0.2) +
  geom_line() +
  facet_grid(rows = vars(variable), labeller= labeller(function(var,value){
  return(parlab[value])})) +
  labs(x="Last Iteration in Chain", y="Shrink Factor",
       colour="Quantile", linetype="Quantile") +
  scale_linetype_manual(values=c(2,1))+
  theme_classic()+
  theme(legend.position = 'top')

ggarrange(traces.gam,gelman.gam)
```

<!-- (ref:diagNeg) MCMC convergence diagnostics for the penalty model.  -->
<!-- On the left trace-plots, on the right changes along the chain of the  -->
<!-- Gelman and Rubin's shrink factor [@brooks_General_1998] -->

<!-- ```{r diag.neg,fig.cap='(ref:diag.neg)'} -->
<!-- longMCMCdata3<-melt(MCMCdata3,id.vars = c("iteration","seed"), -->
<!--                     variable.name = "parameter", -->
<!--                     measure.vars = c("negReward","scaleConst")) -->

<!-- parlab<-list(`negReward` =expression(eta), -->
<!--           `scaleConst`="Scaling constant") -->

<!-- traces.neg<-ggplot(longMCMCdata3,aes(x=iteration,y=value,colour=as.factor(seed)))+ -->
<!--   geom_line()+ -->
<!--   scale_color_manual(values = multDiscrPallet)+ -->
<!--   labs(color="Chain")+ -->
<!--   facet_grid(rows=vars(parameter),scales = "free",switch = "y", -->
<!--              labeller = function(variable,value){ -->
<!--   return(parlab[value])})+ -->
<!--   theme_classic()+ -->
<!--   theme(legend.position = 'top') -->

<!-- pdf(file = NULL) -->
<!-- gp.dat.neg<-gelman.plot(mcmcList.3) -->
<!-- dev.off() -->

<!-- gp.dat.frame.neg <- data.table(rbind(as.data.frame(gp.dat.neg[["shrink"]][,,1]),  -->
<!--                           as.data.frame(gp.dat.neg[["shrink"]][,,2])),  -->
<!--                 q=rep(dimnames(gp.dat.neg[["shrink"]])[[3]], each=nrow(gp.dat.neg[["shrink"]][,,1])), -->
<!--                 last.iter=rep(gp.dat.neg[["last.iter"]], length(gp.dat.neg))) -->

<!-- gelman.neg<-ggplot(melt(gp.dat.frame.neg, c("q","last.iter"), value.name="shrink_factor"),  -->
<!--        aes(last.iter, shrink_factor, colour=q, linetype=q)) +  -->
<!--   geom_hline(yintercept=1, colour="grey30", lwd=0.2) + -->
<!--   geom_line() + -->
<!--   facet_grid(rows = vars(variable), labeller= labeller(function(var,value){ -->
<!--   return(parlab[value])})) + -->
<!--   labs(x="Last Iteration in Chain", y="Shrink Factor", -->
<!--        colour="Quantile", linetype="Quantile") + -->
<!--   scale_linetype_manual(values=c(2,1))+ -->
<!--   theme_classic()+ -->
<!--   theme(legend.position = 'top') -->

<!-- ggarrange(traces.neg,gelman.neg) -->
<!-- ``` -->