### 2022-07-20 manuscript_BE/Title_page.Rmd
New version of the manuscript

---
title: | 
 \large \textbf{Title:} Computational models reveal how cleaner fish adjust decisions in a biological market
subtitle: |
  \normalsize \textbf{Short title:} Cleaner fish cognitive mechanisms in a biological market
author: 
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: a
    corresponding: andresqp@gmail.com 
    fulladress: Rue Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486
  - name: Zegni Triki
    affiliation: 
      - a
      - b
  - name: Redouan Bshary
    affiliation: a
address:
  - code: a
    address: Institute of Biology, University of Neuchatel, Neuchatel, Switzerland
  - code: b
    address: Department of Zoology, Stockholm University, Stockholm, Sweden
output:    
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    template: template.latex
---

## Funding
This work was supported by the Swiss National Science Foundation grants
(grants number: 31003A_153067/1, 310030B_173334/1) to R.B.

## Acknowledgements
ZT and RB kindly thank the staff of Lizard Island Research Station, 
Dominique Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, 
Elena Levorato and William McNeely for their field support. 
AQ thanks Florian Hartig for statistical advice, and the cluster of the 
Institute of Computer Science of the University of Neuchâtel (IIUN) for the
computing resources provided.