### 2022-07-20 manuscript_BE/Title_page.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={ Computational models reveal how cleaner fish adjust decisions in a biological market},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\large \textbf{Title:} Computational models reveal how cleaner
fish adjust decisions in a biological market}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\normalsize \textbf{Short title:} Cleaner fish cognitive
mechanisms in a biological market}
\date{}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\newcommand{\getAff}[1]{
  \getIndex{#1}{a,b}
}

\begin{document}
\maketitle
\begin{flushleft}

Andrés E. Quiñones\textsuperscript{\getAff{a}}\textsuperscript{*},
Zegni Triki\textsuperscript{\getAff{a}, \getAff{b}},
Redouan Bshary\textsuperscript{\getAff{a}}\\
\bigskip
\textbf{\getAff{a}}Institute of Biology, University of Neuchatel,
Neuchatel, Switzerland\\
\textbf{\getAff{b}}Department of Zoology, Stockholm University,
Stockholm, Sweden\\
\bigskip
* Corresponding author: \href{mailto:andresqp@gmail.com}{\nolinkurl{andresqp@gmail.com}}, Rue
Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486\\
\end{flushleft}

\hypertarget{funding}{%
\subsection{Funding}\label{funding}}

This work was supported by the Swiss National Science Foundation grants
(grants number: 31003A\_153067/1, 310030B\_173334/1) to R.B.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

ZT and RB kindly thank the staff of Lizard Island Research Station,
Dominique Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena
Levorato and William McNeely for their field support. AQ thanks Florian
Hartig for statistical advice, and the cluster of the Institute of
Computer Science of the University of Neuchâtel (IIUN) for the computing
resources provided.

\end{document}